ggplot2 practice
=============================
based on http://www.cookbook-r.com/Graphs/

bar & line graphs
-----------------------------

### construct data frame
``` {r}
df <- data.frame(time = factor(c("Lunch", "Dinner"), levels = c("Lunch","Dinner")), total_bill = c(14.89, 17.23))

library(ggplot2)
```

### basic graph

``` {r} 
ggplot(data=df,aes(x=time,y=total_bill)) + geom_bar(stat="identity")
```

ggplot instantiates a ggplot object, aes() specifies aesthetics, geom_bar is a **layer**, necessary to plot the object. specifies what to graph. stat = identity means values in the column of the data frame are plotted. stat = bin would bin them.

### fill colors

``` {r} 
ggplot(data=df,aes(x=time,y=total_bill, fill=time)) + geom_bar(stat="identity")
# equivalently:
ggplot(data=df,aes(x=time,y=total_bill)) + geom_bar(aes(fill=time), stat="identity")
```

fill = the variable you want to color by.

### add outline and remove legend
``` {r}
ggplot(data=df,aes(x=time,y=total_bill, fill=time)) + geom_bar(color="black", stat="identity") + guides(fill=FALSE)
```

geom_bar's color var allows you to set outline color.
another ggplot attribute is guides, pertaining to legend (which I guess is implemented by default and you only need to specify it to remove.. or alter, presumably)

### fancify 

``` {r}
ggplot(data=df, aes(x=time, y=total_bill, fill=time)) + 
  geom_bar(color="black", fille="#DD8888", width=.7, stat="identity") +
  guides(fill=FALSE) +
  xlab("Time of day") + ylab("Total bill") + 
  ggtitle("Average bill for 2 people")
```

neatly organized by category!

### now exploring binning

``` {r}
library(reshape2)
head(tips)

ggplot(data=tips, aes(x=day)) + geom_bar()
```

stat=bin is the default, so you will get COUNTS of your x vars by default. all you need to specify is what your x axis should be

### line graphs

``` {r}
ggplot(data=df, aes(x=time, y=total_bill, group=1)) + geom_line()
#equivalently
ggplot(data=df, aes(x=time, y=total_bill)) + geom_line(aes(group=1))
```

group data so ggplot knows how to connect points. group=1 means all points connected. with more variables and lines, grouping usually done by variable. geom_line() used to specify line graph

### add points

``` {r}
ggplot(data=df, aes(x=time, y=total_bill, group=1)) + geom_line() + geom_point()
```

all that's need is to add geom_point() attribute!

### fancify

``` {r}
ggplot(data=df, aes(x=time,y=total_bill, group=1)) +
  geom_line(color="red", linetype="dotted", size=1.5) +
  geom_point(color="red", size=4, shape=21, fill="white") +
  ylim(0, max(df$total_bill)) + 
  xlab("Time of day") + ylab("Total bill") +
  ggtitle("Average bill for 2 people")
```

### multiple variables

``` {r}
df1 <- data.frame(sex = factor(c("Female","Female","Male","Male")),
                  time = factor(c("Lunch","Dinner","Lunch","Dinner"), levels=c("Lunch","Dinner")),
                  total_bill = c(13.53, 16.81, 16.24, 17.42))
```

### bar graph

``` {r}
ggplot(data=df1, aes(x=time, y=total_bill, fill=sex)) + geom_bar(stat="identity")
```

default is stacked

``` {r}
ggplot(data=df1, aes(x=time, y=total_bill, fill=sex)) + geom_bar(stat="identity", position="dodge", color="black") + scale_fill_manual(values=c("#999999","#E69F00"))
```

position=position_dodge() (or position="dodge"") dodges overlaps to the side. can specify width and heigh here. 

scale\_fill_manual() allows you to specify values to use as colors (ordered).

### flip fill/axis mapping

``` {r}
ggplot(data=df1,aes(x=sex,y=total_bill,fill=time)) + geom_bar(stat="identity",position="dodge")
```

### line graphs

``` {r}
ggplot(data=df1, aes(x=time, y=total_bill, group=sex, color=sex)) + geom_line() + geom_point()

ggplot(data=df1, aes(x=time, y=total_bill, group=sex, size=sex,shape=sex)) +
  geom_line() +
  geom_point(size=3, fill="white") +
  scale_shape_manual(values=c(22,21)) + 
  scale_size_manual(values=c(1,2))
```

can use scale\_blahblahblah_manual to group by a number of attributes -- eg line thickness, point shape, color, probably dash style. easy peasy lemon squeezy

### setting legend titles

` scale_fill_hue(name="title")`

is a way to name legend titles. have to repeat for every scale() you add, so multiple legends are not created.




### graph themes

`theme_bw()` 

adds that theme.

AFTER that line

`theme(legend.position=c(.7, .4))`

positions the legend inside the graph.

I wonder why this goes within theme.. 



theme_gray
The signature ggplot2 theme with a grey background and white gridlines, designed to put the data forward yet make comparisons easy.

theme_bw
The classic dark-on-light ggplot2 theme. May work better for presentations displayed with a projector.

theme_linedraw
A theme with only black lines of various widths on white backgrounds, reminiscent of a line drawings. Serves a purpose similar to theme_bw. Note that this theme has some very thin lines (<< 1 pt) which some journals may refuse.

theme_light
A theme similar to theme_linedraw but with light grey lines and axes, to direct more attention towards the data.

theme_minimal
A minimalistic theme with no background annotations.

theme_classic
A classic-looking theme, with x and y axis lines and no gridlines.


### numeric vals as categorical variables on the x axis

either convert to a factor w/in the dataframe, or `x=factor(var)` directly in the plot

plotting means and error bars
-----------------------------

``` {r init package, load data, summarize}
library(plyr)
source("Desktop/Stanford/Courses/Psy252\ -\ Stats/helpers.R")

df <- ToothGrowth
dfc <- summarySE(df, measurevar = "len", groupvars = c("supp","dose"))
```

### line graphs

``` {r line graphs}
#error bars = SE
ggplot(dfc, aes(x=dose, y=len,color=supp)) + 
  geom_errorbar(aes(ymin=len-se, ymax=len+se), width=.1) +
  geom_line() +
  geom_point()

#fix error bar overlap
pd <- position_dodge(.1)

ggplot(dfc, aes(x=dose, y=len,color=supp)) + 
  geom_errorbar(aes(ymin=len-se, ymax=len+se), width=.1, position=pd) +
  geom_line() +
  geom_point()

#switch to 95% CI
ggplot(dfc, aes(x=dose, y=len,color=supp)) + #add group=supp for black error bars
  geom_errorbar(aes(ymin=len-ci, ymax=len+ci), width=.1, position=pd) + #add color="black" (not in aes())
  geom_line() +
  geom_point()
```

### fancify

``` {r fancify line graph}
ggplot(dfc, aes(x=dose, y=len, color=supp, group=supp)) +
  geom_errorbar(aes(ymin=len-se, ymax=len+se), color="black", width=.1, position=pd) +
  geom_line(position=pd) +
  geom_point(position=pd, size=3, shape=21, fill="white") +
  xlab("Dose (mg)") +
  ylab("Tooth length") +
  scale_color_hue(name="Supplemental type",
                  breaks=c("OJ", "VC"),
                  labels=c("Orange juice","Ascorbic acid"),
                  l=40) + # lightness 40 -> darker colors (across plot)
  ggtitle("The Effecto f Vitamin C on\nTooth Growth in Guinea Pigs") +
  scale_y_continuous(limits=c(0, max(dfc$len + dfc$se)), breaks=0:20*4) +
  theme_bw() +
  theme(legend.justification=c(1,0), legend.position=c(1,0))
```

order of layers matters. draw points on last to be able to fill with white on top of the lines.

### bar graphs

need the size var to be a _factor_

``` {r bar graph}
dfc2 <- dfc
dfc2$dose <- factor(dfc2$dose)

#SE
ggplot(dfc2, aes(x=dose, y=len, fill=supp)) +
  geom_bar(position="dodge", stat="identity") +
  geom_errorbar(aes(ymin=len-se, ymax=len+se),
                width=.2,
                position=position_dodge(.9))

#CI
ggplot(dfc2, aes(x=dose, y=len, fill=supp)) +
  geom_bar(position="dodge", stat="identity") +
  geom_errorbar(aes(ymin=len-ci, ymax=len+ci),
                width=.2,
                position=position_dodge(.9))
```

### within-subjects

tightens error bars. variance treated differently.

###one var

``` {r w/in subj 1 var}
dfw <- read.table(header=T, text='
 subject pretest posttest
       1    59.4     64.5
       2    46.4     52.4
       3    46.0     49.7
       4    49.0     48.7
       5    32.5     37.4
       6    45.2     49.5
       7    60.3     59.9
       8    54.3     54.1
       9    45.4     49.6
      10    38.9     48.5
 ')

dfw$subject <- factor(dfw$subject)

#reshape2 should be loaded

dfw.long <- melt(dfw, id.vars = "subject", measure.vars = c("pretest", "posttest"), variable.name = "condition")

dfwc <- summarySEwithin(dfw.long, measurevar="value", withinvars="condition", idvar="subject", na.rm=FALSE, conf.interval=.95)

# TODO look into summarySEwithin to see how it calculates CI differently for w/in subj
# value and value_norm are means (un/normalized)

ggplot(dfwc, aes(x=condition, y=value, group=1)) +
  geom_line() +
  geom_errorbar(width=.1, aes(ymin=value-ci, ymax=value+ci)) +
  geom_point(shape=21, size=3, fill="white") +
  ylim(40,60)
```

### what would individual differences in condition look like?

``` {r indiv diffs}

ymax <- max(dfw.long$value)
ymin <- min(dfw.long$value)

# plots of vals for individuals
ggplot(dfw.long, aes(x=condition, y=value, color=subject, group=subject)) + geom_line() + geom_point(shape=21, fill="white") + ylim(ymin, ymax)

# normalize data
dfwNorm.long <- normDataWithin(data=dfw.long, idvar="subject", measurevar="value")

# plot norm'd vals for individuals
ggplot(dfwNorm.long, aes(x=condition, y=value_norm, color=subject, group=subject)) + geom_line() + geom_point(shape=21, fill="white") + ylim(ymin, ymax)
```

### differences in error bars -- between-subject and within-subject

``` {r btwn v w/in subj error bars}
dfwc.between <- summarySE(data=dfw.long, measurevar="value", groupvars="condition", na.rm=FALSE, conf.interval=.95)

# plot both
ggplot(dfwc.between, aes(x=condition, y=value, group=1)) + 
  geom_line() +
  geom_errorbar(width=.1, aes(ymin=value-ci, ymax=value+ci), color="red") +
  geom_errorbar(width=.1, aes(ymin=value-ci, ymax=value+ci), data=dfwc) +
  geom_point(shape=21, size=3, fill="white") +
  ylim(ymin, ymax)
```

so you can layer error bars! again, order appears to matter.

### two vars

``` {r two var w/in subj}
data <- read.table(header=T, text='
 Subject RoundMono SquareMono RoundColor SquareColor
       1        41         40         41          37
       2        57         56         56          53
       3        52         53         53          50
       4        49         47         47          47
       5        47         48         48          47
       6        37         34         35          36
       7        47         50         47          46
       8        41         40         38          40
       9        48         47         49          45
      10        37         35         36          35
      11        32         31         31          33
      12        47         42         42          42
')

# convert to long form
data.long <- melt(data=data, id.var="Subject",
                  measure.vars=c("RoundMono", "SquareMono", "RoundColor", "SquareColor"),
                  variable.name="Condition")
names(data.long)[names(data.long)=="value"] <- "Time"

# split cond into shape & colorscheme
data.long$Shape <-NA
data.long$Shape[grepl("^Round", data.long$Condition)] <-"Round"
data.long$Shape[grepl("^Square", data.long$Condition)] <-"Square"
data.long$Shape <- factor(data.long$Shape)

data.long$ColorScheme <- NA
data.long$ColorScheme[grepl("Mono$", data.long$Condition)] <- "Monochromatic"
data.long$ColorScheme[grepl("Color$", data.long$Condition)] <- "Colored"
data.long$ColorScheme <-factor(data.long$ColorScheme, levels=c("Monochromatic","Colored"))

#clean up -- remove Condition col
data.long$Condition <- NULL

data.long

datac <- summarySEwithin(data.long, measurevar="Time", withinvars=c("Shape","ColorScheme"), idvar="Subject")

ggplot(datac, aes(x=Shape, y=Time, fill=ColorScheme)) +
  geom_bar(position=position_dodge(.9), color="black", stat="identity") +
  geom_errorbar(position=position_dodge(.9), width=.25, aes(ymin=Time-ci, ymax=Time+ci)) + 
  coord_cartesian(ylim=c(40,46)) +
  scale_fill_manual(values=c("#CCCCCC", "#FFFFFF")) +
  scale_y_continuous(breaks=seq(1:100)) +
  theme_bw() +
  geom_hline(yintercept=38)
```


