Anna Khazenzon, Homework 2, 10/8/14
=================================
``` {r load libraries}
library(ggplot2)
library(scales)
```

Question 1
---------------------------------
Here, we calculate mean, sum of squares, variance, standard deviation, and standard error for a sample (_N_=3).

``` {r calc central tendency and variance measures}
sample <- c(1, 1, 0)

s_mean <- (sample[1] + sample[2] + sample[3])/(length(sample))
s_ss <- (sample[1] - s_mean)^2 + (sample[2] - s_mean)^2 + (sample[3] - s_mean)^2
s_var <- s_ss/(length(sample) - 1)
s_sd <- sqrt(s_ss/(length(sample) - 1))
s_se <- s_sd/sqrt(length(sample))
  

```
The equation to calculate the sample mean $\bar{x}$ is:

$$\bar{x} = \frac{\sum_{i=1}^{n} x_i}{n}$$

For our sample, the mean $\bar{x}$ is `r s_mean`.

The equation to calculate sum of squares $SS$ is:

$$SS =  \sum_{i=1}^{n} (x_i - \bar{x})^2$$

For our sample, the sum of squares $SS$ is `r s_ss`.

The equation to calculate variance $s^2$ is:

$$s^2 = \frac{SS}{n - 1} $$

For our sample, the variance $s^2$ is `r s_var`.

The equation to calculate sample standard deviation $s$ is:

$$s = \sqrt{\frac{SS}{n - 1}}$$

For our sample, the standard deviation $s$ is `r s_sd`.

The equation to calculate standard error of $\bar{x}$, $SE_\bar{x}$ is:

$$SE_\bar{x} = \frac{s}{\sqrt{n}}$$

For our sample, the standard error of $\bar{x}$, $SE_\bar{x}$ is `r s_se`.

Question 6
---------------------------------
Within the memory bias experiment dataset, we were interested in whether _Past Happiness_ and _Future Happiness_ were differentially related across _Memory Group_ (free, biased, varied). To explore this, we looked at the correlations and plots of _Past Happiness_ and _Future Happiness_ within each _Memory Group_.

``` {r import and recode data}
memdata <- read.csv('http://web.stanford.edu/class/psych252/_downloads/hw2data.csv',header=T, sep=',')

#Type = exper cond, 1=free, 2=biased, 3=varied
#Pasthapp = rating
#Responsible = self-reported feelings of responsibility
#Futurehapp = rating of expected future happiness
#FTP = future time perspective
#complain = yes/no for whether participants considered complaining

# from Q3, recoded variables
# look at data ranges and histograms to determine how to recode
sapply(memdata, range)
hist(memdata$Pasthapp)
hist(memdata$Responsible)
hist(memdata$Futurehapp)
hist(memdata$FTP)

memdata$Type <- factor(memdata$Type)
type_names <- c("1" = "Free", "2" = "Biased", "3"="Varied")

# recoding Pasthapp into Low, Med, and High
memdata$pastcat <- findInterval(memdata$Pasthapp, c(0,5,10))
# recoding Responsible into Low and High
memdata$rescat <- findInterval(memdata$Responsible, c(0,10))
# recoding Futurehapp into Low, Med, and High
memdata$futcat <- findInterval(memdata$Futurehapp, c(0,5,10))
# recoding FTP into Low and High
memdata$ftpcat <- findInterval(memdata$FTP, c(0,10))
```

``` {r rel btwn past and future happ by type - continuous}
Q6_rs <- vector("list", 3)
for (i in levels(memdata$Type)) {
  Q6_rs[[i]] = cor.test(~ Pasthapp + Futurehapp, memdata, Type==i, na.action=na.omit)
  plot(memdata[memdata$Type==i, c(2,4)], type = 'p', main = type_names[i])
  lines(lowess(memdata[memdata$Type==i, c(2,4)]), lty = 2)
}
print(Q6_rs["1"])

print(Q6_rs["2"])

print(Q6_rs["3"])
```

``` {r rel btwn past and future happ by type - categorical}
Q6_rs <- vector("list", 3)
for (i in levels(memdata$Type)) {
  Q6_rs[[i]] = cor.test(~ pastcat + futcat, memdata, Type==i, na.action=na.omit)
  plot(memdata[memdata$Type==i, c(7,9)], type = 'p', main = type_names[i])
  lines(lowess(memdata[memdata$Type==i, c(7,9)]), lty = 2)
}
print(Q6_rs["1"])

print(Q6_rs["2"])

print(Q6_rs["3"])
```

Question 7
---------------------------------
Next, we tested whether _Future Happiness_ depends on _Memory Group_.

``` {r future hap ~ mem group?}
Q7_rs1 <- lm(memdata$Futurehapp ~ memdata$Type)
summary(Q7_rs1)

Q7_rs2 <- oneway.test(memdata$Futurehapp ~ memdata$Type)
print(Q7_rs2)

# boxplot to decide whether or not to transform data
ggplot(memdata, aes(x=Type, y=Futurehapp, fill=Type)) + geom_boxplot() + theme_bw() + guides(fill=FALSE)
# we see that variances appear to be unequal. We test this using the bartlett test for homogeneity of variance.
Q7_rs3 <- with(memdata, bartlett.test(Futurehapp ~ Type, na.action=na.omit))
# we see that the variances are not homogeneous among the three memory groups (p=0.017)

# to meet the assumptions of an ANOVA we should transform our data to have equal variance
memdata$fut_log <- log(memdata$Futurehapp + 0.5)
# we run the bartlett test on the transformed data
Q7_rs4 <- with(memdata, bartlett.test(fut_log ~ Type, na.action=na.omit))
# we see that the variances are still not homogeneous (p=0.00026)
```

Based on both the general linear model and one way ANOVA, it appears as though _Future Happiness_ does not depend on _Memory Group_. The linear model in which _Memory Group_ predicts _Future Happiness_ is better than the one-way ANOVA testing the relationship between the two variables, but neither approach significance.(throw some p values in here)

Based on a boxplot of the data, we could see that variance looked inhomogeneous across _Memory Group_ types. We calculated the homogeneity of variance (bartlett p) We log transformed the data, but this did not decrease the inhomogeneity (bartlett p). 

do some explaining, speculating.

Question 9
---------------------------------
Next, we wanted to establish whether there was a relationship between _Responsible_ and _Complain_, using _logistic regression_. We predicted tht individuals who felt less responsible for the incident were more likely to complain than those who blamed themselves.

``` {r test responsible and complain rel}
#glm(,family = binomial) - just use to assess whether effect of each predictor var on dependent categorical var significant -- see shw2sols.r
Q9_rs_cat <- glm(memdata$complain ~ memdata$rescat, family=binomial)

Q9_rs_cont <- glm(memdata$complain ~ memdata$Responsible, family=binomial)

#plot w/ 2 facets - cat and cont - i haven't done scatter w/ggplot yet huh
#ggplot(aes(x=rescat, y=complain)) +
  
```

# explain relationship

Because we predicted that individuals who felt responsible would not complain, and those who did not feel responsible would, we first ran the generalized linear model using the coded Responsible variable rescat. We found a significant negative effect of responsibility on complaining, p<.01. 

To further investigate this relationship, we ran the model using the full continuous variable Responsible, assessing whether as individuals felt incrementally more or less responsible, they became increasingly less or more likely to complain. We found a significant negative effect of responsibility on complaining in this model as well, p<.01. The relationship was even stronger. This may be the case because partitioning responsibility into high and low meant including mid-range values in either group, watering down the effect.


Question 10
---------------------------------
We now looked at a population known to be Normal, with standard deviation, $\sigma$ = 2. The mean, $\mu$, was unknown. We tested the null hypothesis $H_0: \mu = 4$ against the alternative hypothesis $H_1: \mu = 4.4$, using a sample size, _N_ = 96, and $\alpha$ = 0.05.

To calculate the power of this test, we first calculated the critical value _c_.

``` {r calc power}
# calc c
c <- qnorm(0.05, mean=4.0, sd=2/sqrt(96), lower.tail=F)
power <- pnorm(c, mean=4.4, sd=2/sqrt(96), lower.tail=F)
```

The power of this test was ` r round(power,2)`. (add a line about why it was justified to run this as one-tailed -- only 2 possibilities for mu are 4.0 and 4.4)

Question 12
---------------------------------

We wanted to compare _cognitive behavioral therapy (CBT)_ and _dialectical behavioral therapy (DBT)_ to assess the success of these two intervention techniques. We tested change scores for each group to see which produced more pronounced changes.

``` {r test CBT v DBT change}

CBT <- c(13,21,9,12,9,2,3,6,4,7)
DBT <- c(5,1,2,2,6,0,0,4)

Q12_rs <- t.test(CBT, DBT)
```

It is difficult to interpret our results, given that we only had change scores for the two therapeutic techniques. Because the values were not standardized, we do not know what a change on one measure corresponds to on the other, so although _CBT_ resulted in significantly greater change than _DBT_ based on score alone (t, df, p), this does not mean it was more effective.